---
title: "Reproducible Research Project 1"
author: "nragonese"
date: "December 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
#packagesused
library(tidyverse)
library(ggplot2)

#Be sure to set your preferred working directory
setwd("C:/UserData/OneDrive - Red Ventures/Desktop/Data Science Accelerator/reproducible-research")

# URL with dataset
dataURL <- "https://github.com/nragonese/RepData_PeerAssessment1/blob/master/activity.zip"

#only download if file doesn't exists
if(!file.exists("Activity Data")) {
  dir.create("Activity Data")
  
 fp <- "./Activity Data.zip" #File Path for the data
  
  download.file(dataURL, fp)
  unzip(fp,exdir = WD) #function to unzip file
  file.remove(fp) #Removes the zip file because we've already imported the data
}


#Read files into R
ActivityData <- read.csv("activity.csv")


```

### What is the mean total number of steps taken per day?
10,766 is the mean total number of steps taken per day, 10,765 is the media number of steps taken.

```{r}
byday <- setNames(aggregate(ActivityData$steps, by= list(date = ActivityData$date), FUN =sum),c("date","steps"))

hist(byday$steps, breaks = 50, col = "red", xlab = "Daily Total Steps", main = "Histogram of Daily Step Totals")

summary(byday$steps)
```

### What is the average daily activity pattern?

```{r}

byinterval <- setNames(aggregate(ActivityData$steps, by = list(interval = ActivityData$interval), FUN =mean, na.rm = TRUE),c("interval","steps"))

plot(byinterval$steps ~ byinterval$interval, type = "l", xlab = "Interval", ylab = "Steps", main = "Average Steps per Interval")

subset(byinterval,byinterval$steps == max(byinterval$steps))

```

Interval 835 has the most steps on average with about 206.

### Imputing missing values 


```{r}
sum(is.na(ActivityData$steps))
```

There are 2304 missing entries.These missing entries will be replaced with the average of the interval when the missing value occured.

```{r}
merged <- merge(ActivityData, byinterval, by = "interval")

merged$steps.x[is.na(merged$steps.x)]= merged$steps.y[is.na(merged$steps.x)]

mergedbyday <- setNames(aggregate(merged$steps.x, by= list(date = merged$date), FUN =sum),c("date","steps"))

hist(mergedbyday$steps, breaks = 50, col = "red", xlab = "Daily Total Steps", main = "Histogram of Daily Step Totals")

summary(mergedbyday$steps)
```

Both the mean and the median remained almost identical, while there was a shift from the 3rd quartile to the 1st quartile.

### Are there differences in activity patterns between weekdays and weekends?

```{r}

merged$weekday <- ifelse(weekdays(as.Date(ActivityData$date)) %in% c("Saturday","Sunday"),"weekend","weekday")

mergedbydaytype <- setNames(aggregate(merged$steps.x, by = list(merged$weekday, merged$interval), FUN = mean),c("weekday","interval","steps"))

library(lattice)
xyplot(steps ~ interval | weekday, mergedbydaytype, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Mean Steps")
```

